---
title: Network Plots and Basic Statistics
author: Fan Yang
date: July 11, 2017
---

```{r global_options, include=FALSE}
library(knitr)
library(rmarkdown)
opts_knit$set(root.dir='/Users/fanyang/Box Sync/Mexico/maurilia/data_for_R/')
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# Objecttive:
Visually interprete the co-occurrence analsis and the basic statistics of the network plots. 

# Recall:
I previously introduced the co-occurrence analysis in R [here](https://github.com/germs-lab/tutorials/blob/master/2016-06-15_cooccurrence/spearman_hoeffding_tutorial.md). 
Therefore, I will start with generating a simulated co-occurrence relationship. 

# Procedures:
## 1. Load libraries:
```{r}
library(Hmisc)
library(reshape2)
library(igraph)
library(sna)
```

## 1. Simulating a set of data for 4 types of relationship: 
```{r}
set.seed(35)
x <- rnorm(100, mean=1, sd=9)
independent<-0+rnorm(100, mean=14, sd=9)
linear <- rnorm(100, mean=17, sd=5) +0.8*x
monotonic <- rnorm(100, mean=800, sd=1000) + x^3
non.mono <-  rnorm(100, mean=100, sd=100) - x^2
#add some sample ID and information:
id <- paste("S", seq(1, 100), sep="_")
type <- rep(c("A", "B", "C", "D", "E"), 20)

#put them all together
df<-data.frame(id, type, x, independent, linear, monotonic, non.mono)
head(df)
```

## 2. Evaluate their relationships:
```{r}
results_sp<-rcorr(as.matrix(df[, 3:7]),type="spearman")
results_hd<-hoeffd(as.matrix(df[, 3:7])) 

#make two seperate objects for p-value and correlation coefficients
rhos<-results_sp$r
sp_ps<-results_sp$P
ds<-results_hd$D
ds_ps<-results_hd$P

# going to melt these objects to 'long form' where the first two columns make up the pairs of OTUs, I am also removing NA's as they are self-comparisons, not enough data, other bad stuff
sp_melt<-na.omit(melt(sp_ps))
ds_melt<-na.omit(melt(ds_ps))

#creating a qvalue (adjusted pvalue) based on FDR
sp_melt$spearman_qval<-p.adjust(sp_melt$value, "fdr")
ds_melt$hoeffding_qval<-p.adjust(ds_melt$value, "fdr")
#making column names more relevant
names(sp_melt)[3]<-"spearman_pval"
names(ds_melt)[3]<-"hoeffding_pval"

# now melting the correlations
rho.melt<-na.omit(melt(rhos))
d.melt<-na.omit(melt(ds))
# add column names
names(rho.melt)[3]<-"rho"
names(d.melt)[3]<-"D"

#merging together  then subset
sp_merged<-merge(sp_melt,rho.melt,by=c("Var1","Var2"))
ds_merged<-merge(ds_melt, d.melt,by=c("Var1","Var2"))
merged<-merge(sp_merged, ds_merged, by=c("Var1", "Var2"))

#subset for significant relationships
final_results<-subset(merged, spearman_qval < 0.05 | spearman_qval >= 0.05 & hoeffding_qval < 0.05 | is.na(spearman_qval) | is.na(hoeffding_qval))

head(merged)
```


