---
title: Network Plots and Basic Statistics
author: Fan Yang
date: July 11, 2017
---

```{r global_options, include=FALSE}
library(knitr)
library(rmarkdown)
opts_knit$set(root.dir='~/Documents/repos/tutorials/2017-07-10_network_analysis')
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# Objecttive:
Visually interprete the co-occurrence analsis and the basic statistics of the network plots. 

# Recall:
I previously introduced the co-occurrence analysis in R [here](https://github.com/germs-lab/tutorials/blob/master/2016-06-15_cooccurrence/spearman_hoeffding_tutorial.md). 
Therefore, I will start with generating a simulated co-occurrence relationship. 

# Procedures:
## 1. Load libraries:
```{r}
library(Hmisc)
library(reshape2)
library(igraph)
library(sna)
library(RColorBrewer)
```

## 2. Simulating a set of data for 4 types of relationship: 
```{r}
set.seed(35)
x <- rnorm(100, mean=1, sd=9)
independent<-0+rnorm(100, mean=14, sd=9)
linear <- rnorm(100, mean=17, sd=5) +0.8*x
monotonic <- rnorm(100, mean=800, sd=1000) + x^3
non.mono <-  rnorm(100, mean=1000, sd=20) - x^2

#put them all together
df<-data.frame(x, independent, linear, monotonic, non.mono)
head(df)
```

### 2.1. random relationship
```{r}
plot(independent ~ x, data = df)
```

### 2.2. linear relationship
```{r}
plot(linear ~ x, data = df)
```

### 2.3. monotonic relationship
```{r}
plot(monotonic ~ x, data = df)
```

### 2.4. non-monotonic relationship
```{r}
plot(non.mono ~ x, data = df)
```

## 3. Evaluate their relationships:
```{r}
results_sp<-rcorr(as.matrix(df),type="spearman")
results_hd<-hoeffd(as.matrix(df)) 

#make two seperate objects for p-value and correlation coefficients
rhos<-results_sp$r
sp_ps<-results_sp$P
ds<-results_hd$D
ds_ps<-results_hd$P

# going to melt these objects to 'long form' where the first two columns make up the pairs of OTUs, I am also removing NA's as they are self-comparisons, not enough data, other bad stuff
sp_melt<-na.omit(melt(sp_ps))
ds_melt<-na.omit(melt(ds_ps))

#creating a qvalue (adjusted pvalue) based on FDR
sp_melt$spearman_qval<-p.adjust(sp_melt$value, "fdr")
ds_melt$hoeffding_qval<-p.adjust(ds_melt$value, "fdr")
#making column names more relevant
names(sp_melt)[3]<-"spearman_pval"
names(ds_melt)[3]<-"hoeffding_pval"

# now melting the correlations
rho.melt<-na.omit(melt(rhos))
d.melt<-na.omit(melt(ds))
# add column names
names(rho.melt)[3]<-"rho"
names(d.melt)[3]<-"D"

#merging and subset for significant coefficients
sp_merged<-merge(sp_melt,rho.melt,by=c("Var1","Var2"))
ds_merged<-merge(ds_melt, d.melt,by=c("Var1","Var2"))
merged<-merge(sp_merged, ds_merged, by=c("Var1", "Var2"))
#subset for significant relationships
merged<-subset(merged, spearman_qval < 0.05 | spearman_qval >= 0.05 & hoeffding_qval < 0.05 | !is.na(spearman_qval) | !is.na(hoeffding_qval))

#because both rho and D are coefficient, we can combine them
merged$coefficient_type <- ifelse(abs(merged$rho) > abs(merged$D), "rho", "D")
merged$coefficient <- ifelse(merged$coefficient_type == colnames(merged)[5], merged$rho, merged$D)

#remove 
final_results <- merged[, c("Var1", "Var2", "coefficient", "coefficient_type")]
head(final_results)
```

## 4. Network graphs
### 4.1. the edges: 
#### 4.1.1. the base:
```{r}
# turn relationships into a network graph
# the relationships here are undirected
temp.graph<-graph.edgelist(as.matrix(final_results[,c(1,2)]),directed=FALSE)
# E() assigns edge atributes
E(temp.graph)
E(temp.graph)$weight<-abs(final_results$coefficient)
```

#### 4.1.2. the most basic plot:
```{r}
plot(temp.graph)
# add some graph attributes
plot(temp.graph, vertex.size = 5, edge.width=E(temp.graph)$weight)
```

#### 4.1.3. the simplified plot:
```{r}
# because our network is undirected, every relationship is duplicated. "simplify" removes duplicated nodes and edges
temp.simplified<-simplify(temp.graph, edge.attr.comb="mean")
    # by default, it would use "sum"
plot(temp.simplified, edge.width=E(temp.simplified)$weight)
```

#### 4.1.4. we can exaggerate the line thickness:
```{r}
E(temp.simplified)$weight <- E(temp.simplified)$weight * 10
plot(temp.simplified, edge.width=E(temp.simplified)$weight)
```

#### 4.1.5. and reposition the nodes
```{r}
plot(temp.simplified, edge.width=E(temp.simplified)$weight, layout=layout_in_circle)
plot(temp.simplified, edge.width=E(temp.simplified)$weight, layout=layout_as_tree)
plot(temp.simplified, edge.width=E(temp.simplified)$weight, layout=layout_nicely)
# more here: http://igraph.org/r/doc/layout_.html
```

#### 4.1.6. we can assign different line styles to positive and negative coefficients:
```{r}
E(temp.graph)$coefficient <- final_results$coefficient
E(temp.graph)$lty <- ifelse(E(temp.graph)$coefficient > 0, 1, 2)
temp.simplified<-simplify(temp.graph, edge.attr.comb="mean")
E(temp.simplified)$lty <- ifelse(E(temp.simplified)$coefficient > 0, 1, 2)
plot(temp.simplified, edge.width=E(temp.simplified)$weight, layout=layout_in_circle, edge.lty = E(temp.simplified)$lty)
```

#### 4.1.7. we can assign different colors to positive and negative coefficients as well:
```{r}
E(temp.simplified)$colors <- ifelse(E(temp.simplified)$coefficient > 0, "black", "red")
plot(temp.simplified, edge.width=E(temp.simplified)$weight, layout=layout_in_circle, edge.lty = E(temp.simplified)$lty, edge.color = E(temp.simplified)$colors)
```

### 4.2. the vertexes (nodes)
similar to the edges, we can use `ifelse` to directly assign attributes to the nodes. But a lot of times, it's much easier to assigned attributes from a table format.
#### 4.2.1. get the names and their corresponding order:
```{r}
df <- data.frame(V(temp.simplified)$name)
df
```

#### 4.2.2. we can use `RColorBrewer` to generate colors:
```{r}
colorCount <- length(unique(df[, 1]))
getPalette <- colorRampPalette(brewer.pal(8, "Dark2"))
df$colors <- getPalette(colorCount)
V(temp.simplified)$colors <- df$colors
plot(temp.simplified, edge.width=E(temp.simplified)$weight, layout=layout_in_circle, edge.lty = E(temp.simplified)$lty, edge.color = E(temp.simplified)$colors, vertex.color = V(temp.simplified)$colors)
```

#### 4.2.3. we can remove the node labels and add a legend:
```{r}
plot(temp.simplified, edge.width=E(temp.simplified)$weight, layout=layout_in_circle, edge.lty = E(temp.simplified)$lty, edge.color = E(temp.simplified)$colors, vertex.color = V(temp.simplified)$colors, vertex.label=NA)
legend(x=-1.5, y=-1.1, df[, 1], pch=21, col="black", pt.bg=df$colors, pt.cex=2, cex=.8, bty="n", ncol=1)
```

#### 4.2.4. we can update the node boarder color:
```{r}
V(temp.simplified)$frame.color <- df$colors
plot(temp.simplified, edge.width=E(temp.simplified)$weight, layout=layout_in_circle, edge.lty = E(temp.simplified)$lty, edge.color = E(temp.simplified)$colors, vertex.color = V(temp.simplified)$colors, vertex.frame.color = V(temp.simplified)$frame.color, vertex.label=NA)
legend(x=-1.5, y=-1.1, df[, 1], pch=21, col=df$colors, pt.bg=df$colors, pt.cex=2, cex=.8, bty="n", ncol=1)
```

## 5. Basic network stats:
### 5.1. number of nodes and edges:
```{r}
# number of nodes
N_nodes<-vcount(temp.graph)
N_nodes
# number of edges
N_edges<-ecount(temp.graph)
N_edges
```

### 5.2. how well do our network cluster?
```{r}
# number of clusters
g.components <- clusters(temp.graph)
N_clusters<-g.components$no
N_clusters

# size of the largest cluster in the network
Max_csize<-max(g.components$csize)
Max_csize

# and the number of edges associated with
Max_c_edges<-ecount(induced.subgraph(temp.graph, which(g.components$membership == which.max(g.components$csize))))
Max_c_edges
```

### 5.3. density
the number of edges in a network divided by the total number of possible edges
```{r}
g.density<-graph.density(temp.graph)
g.density
```

### 5.4. the paths around the network on average
```{r}
g.pathlength.avg<-average.path.length(temp.graph)
g.pathlength.avg
```

### 5.5. centrality: 
how connected is the network?

#### 5.5.1. by betweenness: the number of shortest paths from all nodes to all others that pass through that node.
```{r}
betcent<-centralization.betweenness(temp.graph)$centralization
betcent
```

#### 5.5.2. by degree: the number of edges a node has.
```{r}
degcent<-centralization.degree(temp.graph)$centralization
degcent
```

### 5.6.  modularity: 
are nodes connected within their clusters (modules) more than between clusters than they would by chance?
```[r}
g.mod<-modularity(edge.betweenness.community(temp.graph))
g.mod
```

### 5.7. effective size: 
the number of relationships between a node and other unrelated nodes
```{r}
## effective size function ##
# definite a function like this
ego.effective.size <- function(g, ego, ...) {
  egonet <- induced.subgraph(g, neighbors(g, ego, ...))
  n <- vcount(egonet)
  t <- ecount(egonet)
  return(n - (2 * t) / n)
}
effective.size <- function(g, ego=NULL, ...) {
  if(!is.null(ego)) {
    return(ego.effective.size(g, ego, ...))
  }
  return (sapply(V(g), function(x) {ego.effective.size(g,x, ...)}))
}
# calculate average effective size
efsize_avg<-mean(effective.size(temp.graph, mode = "all"))
efsize_avg
```

### 5.8. combine them together
```{r}
test<-cbind(N_nodes, N_edges, N_clusters, Max_csize, Max_c_edges, g.density, g.pathlength.avg, betcent, degcent, g.mod, efsize_avg)
test
```

